---
title: "data_manipulation"
output: github_document
author: Guojing Wu
data: 2018-09-20
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
 
## import data

```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data_import_examples/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## select, remove and rename

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
select(litters_data, group:gd_of_birth)
select(litters_data, -pups_survive)
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

or you can just use rename
```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

choose a specific pattern

```{r}
select(litters_data, starts_with("gd"))
```

## Learning assessment
```{r}
select(pups_data, litter_number, sex, pd_ears)
```

## filter

compare

```{r}
filter(litters_data, gd_of_birth == 20)
filter(litters_data, pups_born_alive >= 2)
```

two ways of unequal

```{r}
filter(litters_data, pups_survive != 4)
filter(litters_data, !(pups_survive == 4))
```

two ways of multi-choose

```{r}
filter(litters_data, group %in% c("Low7", "Low8"))
#filter(litters_data, group == "Low7" | group == "Low8"))
```

## mutate

give an extra column for convenient analyze

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
)
```

## arrange

you can arrange data by one or more column

```{r}
head(arrange(litters_data, group), 10)
head(arrange(litters_data, group, pups_born_alive), 10)
```

## Pipe!!!

the first choice

```{r}
litters_data_raw = read_csv("./data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars
```

the second choice

```{r}
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data_import_examples/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean
```

But both above is cunfusing, we then use piping instead:

```{r}
litters_data = 
  read_csv("./data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>%
  filter(!is.na(gd0_weight))

head(litters_data, 10)
```

illustrate by using dot

```{r}
litters_data = 
  read_csv("./data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(.) %>%
  select(., -pups_survive) %>%
  mutate(., 
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)
  ) %>%
  filter(., !is.na(gd0_weight))

head(litters_data, 10)
```